---
title: "Alluvial plot"
---

```{r}
#| label: setup
#| include: true
pacman::p_load(
               DT,
               dtplyr,
               here,
               knitr,
               tidyverse,
               patchwork,
               metafor,
               orchaRd, ggalluvial, patchwork, ggrepel
               )
```




The alluvial plots below summarize the distribution of effect sizes across key study characteristics (e.g., outcome type, species/strain, sex, and exposure design variables).

::: callout-note
### Post-processing for the manuscript figure

The code here reproduces the alluvial plots and exports them as PDF/JPG.  
For the final manuscript figure, we **made minor edits in vector-graphics software** (e.g., Illustrator/Inkscape) to improve readability—primarily **manual repositioning of labels, spacing adjustments, and small layout tweaks**. These edits **do not change the underlying data or flows**; they only improve figure legibility.
::: 




```{r}
db <- readr::read_csv(here(
  "..","data","db_effect_sizes.csv")) %>%
mutate(across(c(C_n, Ex_n, C_mean, Ex_mean, C_SE, Ex_SE, C_SD, Ex_SD), as.numeric))%>%
  mutate(across(where(is.character), as.factor))
```

```{r}
db_freq <- db %>%
  # Group by all variables that define a unique flow path and the frequency count
  group_by(Study_ID, Outcome_type, Species_latin, Strain, Sex) %>%
  # Count the number of rows (observations) in each group
  summarise(Freq = n(), .groups = 'drop')
```

```{r}
# Shared label size so every text element in the figure (strata names and
# axis headings, across both panels) renders at the same font size.
label_size <- 3.2

# Colour-blind-safe palette (Okabe-Ito). "Mixed" uses a saturated green
# instead of light gray so it stays visible against the white background
# once the geom_flow() alpha is applied.
custom_colors <- c(
  "Female" = "#0072B2",   # Blue
  "Male" = "#D55E00",     # Vermillion
  "Mixed" = "#009E73"     # Bluish green
)

# --- Plotting the Aggregated Data (db_freq) ---
Out_spp_strain<-ggplot(db_freq,
       # Explicitly use the calculated Freq column for the flow height
       aes(y = Freq,
           axis1 = Outcome_type, axis2 = Species_latin, axis3 = Strain)) +

  # 1. Flow and Transparency
  geom_flow(color = "black",
            aes(fill = as.factor(Sex)),
            width = 0.45,
            alpha = 0.45) +

  # Add the strata boxes (widened so long names have more room)
  geom_stratum(width = 0.45) +

  # 2. Adjust Strata Text (Labels) - repelled apart + leader lines so
  # low-frequency (thin) strata don't collide with their neighbours
  geom_text_repel(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = label_size,
            discern = TRUE,
            color = "black",
            fontface = "bold",
            hjust = 0.5,
            vjust = 0.5,
            direction = "y",
            box.padding = 0.35,
            point.padding = 0.15,
            force = 3,
            min.segment.length = 0,
            segment.size = 0.3,
            segment.color = "grey40",
            max.overlaps = Inf,
            seed = 1) +

  # 3. Apply Custom Color Palette
  scale_fill_manual(
    name = "Sex",
    values = custom_colors
  ) +
labs(
    title = "Study subjects")+
  # 4. Axis Headings as real (centred) axis text, so they sit in their own
  # margin band below the panel instead of overlapping the flows
  scale_x_continuous(breaks = 1:3, labels = c("Outcome Type", "Species", "Strain")) +

  # 5. Hide y-Axis and Clean Theme
  scale_y_continuous(labels = NULL) +
  # 6. Legend under the panel, widened so the swatches read easily
  guides(fill = guide_legend(title.position = "top", label.position = "bottom", nrow = 1)) +
  theme(
    plot.margin = margin(t = 5, r = 5, b = 5, l = 5, unit = "pt"),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = label_size * 2.845, margin = margin(t = 10)),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1.4, "cm"),
    legend.key.height = unit(0.4, "cm"),
    legend.title = element_text(face = "bold", size = label_size * 3),
    legend.text = element_text(size = label_size * 3)
    )
```

```{r}
Out_spp_strain
```

```{r}

db_fixed_assay <- db %>%
  # Filter the data to ensure we only rename the categories we listed
  mutate(
    # Create a new, shortened column for the Assay Type labels
    Assay_initials = case_when(
      Assay_type == "Elevated Plus Maze" ~ "EPM",
      Assay_type == "Open Field Test" ~ "OFT",
      Assay_type == "Light-Dark Box" ~ "LDB",
      Assay_type == "Sucrose Preference Test (SPT)" ~ "SPT",
      Assay_type == "Forced Swim Test (FST)" ~ "FST",
      Assay_type == "Tail Suspension Test (TST)" ~ "TST",
      # NSF, Nest building, and Marble burying are each individually too
      # rare to label without colliding; collapse them into one category
      Assay_type == "Novelty Suppressed Feeding (NSF)" ~ "Other",
      Assay_type == "Nest building" ~ "Other",
      Assay_type == "Marble burying" ~ "Other",
      TRUE ~ Assay_type # Keep any other unlisted Assay_type names as they are
    ),
    # Order Age chronologically (developmental stage), with the
    # non-chronological categories ("Mixed", "Unclear") placed last
    Lifestage_exposure = factor(Lifestage_exposure,
      levels = c("Juvenile", "Adolescent", "Young adult", "Adult", "Mixed", "Unclear")),
    # Order Duration from shortest to longest exposure
    Music_exposure_duration = factor(Music_exposure_duration,
      levels = c("Acute", "Short", "Medium", "Long"))
  ) %>%
  # Group by the new axis variables and the flow variable (`Induced behaviour`)
  group_by(Study_ID, Lifestage_exposure, Music_exposure_duration, Assay_initials, `Induced behaviour`) %>%
  # Count the number of observations (Freq)
  summarise(Freq = n(), .groups = 'drop')

# --- 2. PLOTTING STEP ---
custom_colors <- c(
  "Innate" = "blue",
  "Induced" = "red"
)

assay_plot <- ggplot(db_fixed_assay,
       # Use Freq for flow height
       # Axis 3 is now Assay_initials
       aes(y = Freq,
           axis1 = Lifestage_exposure, axis2 = Music_exposure_duration, axis3 = Assay_initials)) +

  # 1. Flow and Transparency
  geom_flow(color = "black",
            aes(fill = as.factor(`Induced behaviour`)),
            width = 0.45,
            alpha = 0.45) +

  # Add the strata boxes (widened so long names have more room)
  geom_stratum(width = 0.45) +

  # 2. Adjust Strata Text (Labels) - Will now show the initials (EPM, SPT, etc.)
  # repelled apart + leader lines so crowded low-frequency strata (e.g. MB,
  # NSF, Nest) don't overlap each other
  geom_text_repel(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = label_size,
            discern = TRUE,
            color = "black",
            fontface = "bold",
            hjust = 0.5,
            vjust = 0.5,
            direction = "y",
            box.padding = 0.35,
            point.padding = 0.15,
            force = 3,
            min.segment.length = 0,
            segment.size = 0.3,
            segment.color = "grey40",
            max.overlaps = Inf,
            seed = 1) +

  # 3. Apply Custom Color Palette and Legend Title
  scale_fill_manual(
    name = "Behavioral Mechanisms",
    values = custom_colors
  ) +
  labs(
    title = "Music exposure"
  ) +

  # 4. Axis Headings as real (centred) axis text, so they sit in their own
  # margin band below the panel instead of overlapping the flows
  scale_x_continuous(breaks = 1:3, labels = c("Age", "Duration", "Assay Type")) +

  # 5. Hide y-Axis and Clean Theme
  scale_y_continuous(labels = NULL) +
  # 6. Legend under the panel, widened so the swatches read easily
  guides(fill = guide_legend(title.position = "top", label.position = "bottom", nrow = 1)) +
  theme(
    plot.margin = margin(t = 5, r = 5, b = 5, l = 5, unit = "pt"),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = label_size * 2.845, margin = margin(t = 10)),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(1.4, "cm"),
    legend.key.height = unit(0.4, "cm"),
    legend.title = element_text(face = "bold", size = label_size * 3),
    legend.text = element_text(size = label_size * 3)
    )
```

```{r}
assay_plot
```

```{r}
plot1<-(Out_spp_strain/assay_plot)+ plot_annotation(
    tag_levels = 'A'
  )
```

```{r}

# Overlap among low-frequency strata (e.g. MB/NSF/Nest, "Both") is now
# handled by geom_text_repel() pushing labels apart with leader lines, so
# the canvas no longer needs to be oversized just to make room for them.
#ggsave(filename = here("..","Plots", "alluvial_raw.pdf"),
      # plot = plot1,
       #dpi = 300, device = cairo_pdf, width = 11,
       #height = 13)


#ggsave(filename = here("..","Plots", "alluvial_raw.jpg"),
       plot = plot1,
       width = 11,
       height = 13,
       dpi = 300)
```









::: callout-note
```{r}
sessionInfo()
```

:::






