---
title: "Sensitivity analysis to the assumed within-study correlation"
---

The overall-effect and moderator models (see `overall_effect.qmd` and `uni_moderators.qmd`) build a cohort-level variance–covariance (VCV) matrix with `metafor::vcalc()`, assuming a within-cohort correlation of `rho = 0.5` between sampling errors of different effect sizes measured on the same cohort of animals (`Cohort_ID`), whenever the true within-cohort correlation is not reported by the primary study. A separate but numerically identical value, `ri = 0.5`, is used inside `calculate_lnRR()`, `calculate_lnVR()`, and `calculate_lnCVR()` (see `effect_size_calculation.qmd`) as the assumed within-subject correlation for the sampling-variance formulas of *dependent* (paired / repeated-measures) comparisons (`Comparison_structure == "Dependent"`, `metafor::escalc(measure = "ROMC"/"VRC"/"CVRC", ri = 0.5)`). These are two conceptually distinct correlation parameters that happen to share the same assumed value of 0.5 by convention:

1.  **Cross-effect-size correlation** (`vcalc(rho = ...)`): the assumed correlation between the sampling errors of *different* effect sizes drawn from the same cohort of animals (e.g., two behavioural outcomes measured on the same group). This affects the off-diagonal covariances of the VCV matrix `V` used in every `rma.mv()` call, for any cohort contributing more than one effect size (57 of the dataset's cohorts contribute more than one outcome).
2.  **Within-subject/within-comparison correlation** (`escalc(ri = ...)`): the assumed correlation between the treatment and control measurements of a *single* dependent (paired/repeated-measures) comparison. This affects only the sampling-variance formula of that one effect size, and in the current database applies to the 40 effect sizes (all from a single study, `Sampaio_2017_PSYNEURO`, a within-subject design measured at four life stages) coded `Comparison_structure == "Dependent"`.

This appendix runs a sensitivity analysis on both parameters. It first varies parameter (1), the cross-effect-size correlation used to build the VCV matrix for the overall `lnRR` and `lnVR` models, because that is the correlation that structures the variance–covariance matrix referenced in the manuscript and by reviewers. It then (further down, in "Recomputing sampling variances across the assumed within-subject correlation") varies parameter (2) as well, by recomputing the sampling variances of the 40 `Comparison_structure == "Dependent"` rows directly from the raw extraction database (`data/db251124.csv`) across a range of assumed values of `ri`, and finally crosses both parameters in a joint `ri` x `rho` grid.

```{r}
#| label: setup
#| include: false
pacman::p_load(
               DT,
               dtplyr,
               here,
               knitr,
               tidyverse,
               patchwork,
               metafor,
               orchaRd,
               kableExtra
               )
```

```{r}
#| label: load_effectsize_db
#| message: false
#| warning: false

db <- readr::read_csv(here("..","data","db_effect_sizes.csv")) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(c(lnRR, lnRR_var, lnVR, lnVR_var, lnCVR, lnCVR_var), as.numeric))
```

## Refitting the overall models across a range of assumed correlations

We refit the same intercept-only multilevel models used in `overall_effect.qmd` (random intercepts for `Study_ID`, `ES_ID`, and `Strain`; REML; `test = "t"`), varying only the `rho` argument passed to `vcalc()` when constructing the VCV matrix. We evaluate `rho = 0.00, 0.25, 0.50, 0.75, 0.90`, spanning independence (`rho = 0`) to a strong assumed within-cohort correlation (`rho = 0.90`).

```{r}
#| label: fit_overall_mv_rho
#| message: false
#| warning: false

rho_values <- c(0.00, 0.25, 0.50, 0.75, 0.90)

fit_overall_mv <- function(dat, yi, vi, rho) {

  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]))

  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2$Cohort_ID,
    obs = dat2$ES_ID,
    rho = rho,
    data = dat2
  )

  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | Study_ID,
      ~1 | ES_ID,
      ~1 | Strain
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )

  list(data = dat2, VCV = VCV, model = mod)
}

overall_sensitivity <- purrr::map_dfr(rho_values, function(r) {

  fit_rr <- fit_overall_mv(db, "lnRR", "lnRR_var", r)
  fit_vr <- fit_overall_mv(db, "lnVR", "lnVR_var", r)

  i2_rr <- orchaRd::i2_ml(fit_rr$model)
  i2_vr <- orchaRd::i2_ml(fit_vr$model)

  dplyr::bind_rows(
    tibble::tibble(
      metric        = "lnRR",
      rho           = r,
      k             = fit_rr$model$k,
      estimate      = as.numeric(coef(fit_rr$model)),
      se            = fit_rr$model$se,
      ci.lb         = fit_rr$model$ci.lb,
      ci.ub         = fit_rr$model$ci.ub,
      tval          = fit_rr$model$zval,
      pval          = fit_rr$model$pval,
      sigma2_Study  = fit_rr$model$sigma2[1],
      sigma2_ES     = fit_rr$model$sigma2[2],
      sigma2_Strain = fit_rr$model$sigma2[3],
      I2_Total      = i2_rr[["I2_Total"]],
      I2_Study_ID   = i2_rr[["I2_Study_ID"]],
      I2_ES_ID      = i2_rr[["I2_ES_ID"]],
      I2_Strain     = i2_rr[["I2_Strain"]]
    ),
    tibble::tibble(
      metric        = "lnVR",
      rho           = r,
      k             = fit_vr$model$k,
      estimate      = as.numeric(coef(fit_vr$model)),
      se            = fit_vr$model$se,
      ci.lb         = fit_vr$model$ci.lb,
      ci.ub         = fit_vr$model$ci.ub,
      tval          = fit_vr$model$zval,
      pval          = fit_vr$model$pval,
      sigma2_Study  = fit_vr$model$sigma2[1],
      sigma2_ES     = fit_vr$model$sigma2[2],
      sigma2_Strain = fit_vr$model$sigma2[3],
      I2_Total      = i2_vr[["I2_Total"]],
      I2_Study_ID   = i2_vr[["I2_Study_ID"]],
      I2_ES_ID      = i2_vr[["I2_ES_ID"]],
      I2_Strain     = i2_vr[["I2_Strain"]]
    )
  )
})
```

```{r}
#| label: tbl_overall_sensitivity
#| tbl-cap: "Overall lnRR and lnVR pooled estimates across assumed within-cohort correlations (rho)."

overall_sensitivity %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed")) %>%
  scroll_box(width = "100%")
```

```{r}
#| label: save_overall_sensitivity_table
readr::write_csv(overall_sensitivity, here("..","data","sensitivity_r_overall_results.csv"))
```

## Figure: overall estimate and 95% CI across rho

```{r}
#| label: fig_overall_sensitivity
#| fig-width: 9
#| fig-height: 5

p_sensitivity <- overall_sensitivity %>%
  mutate(metric = factor(metric, levels = c("lnRR", "lnVR"))) %>%
  ggplot(aes(x = factor(rho), y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci.lb, ymax = ci.ub), size = 0.6) +
  facet_wrap(~ metric, scales = "free_y") +
  labs(
    x = expression(paste("Assumed within-cohort correlation, ", rho)),
    y = "Pooled estimate (95% CI)",
    title = "Overall lnRR and lnVR estimates across assumed values of the VCV correlation"
  ) +
  theme_bw(base_size = 13)

p_sensitivity
```

```{r}
#| label: save-figure_sensitivity
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_r_overall.pdf"),
       plot = p_sensitivity,
       dpi = 300, device = cairo_pdf, width = 9, height = 5)

ggsave(filename = here("..","Plots", "sensitivity_r_overall.jpg"),
       plot = p_sensitivity,
       width = 9, height = 5, dpi = 300)
```

## Anxiety-versus-depression lnVR contrast across rho

The manuscript also reports a depression-minus-anxiety contrast for `lnVR` (see `uni_moderators.qmd`, "Outcome type" section). We check whether that contrast, and the two subgroup pooled estimates it is built from, are sensitive to the assumed correlation. We reuse the same `k >= 5` level-filtering rule applied in `uni_moderators.qmd` (the `both` level of `Outcome_type`, k = 3, is dropped).

```{r}
#| label: contrast_sensitivity
#| message: false
#| warning: false

db_vr <- db %>% filter(!is.na(lnVR), !is.na(lnVR_var))

k_counts    <- db_vr %>% count(Outcome_type)
keep_levels <- k_counts %>% filter(n >= 5) %>% pull(Outcome_type)

dat_f <- db_vr %>%
  filter(Outcome_type %in% keep_levels) %>%
  mutate(Outcome_type = droplevels(Outcome_type))

contrast_sensitivity <- purrr::map_dfr(rho_values, function(r) {

  VCV <- metafor::vcalc(
    vi = dat_f$lnVR_var, cluster = dat_f$Cohort_ID, obs = dat_f$ES_ID,
    rho = r, data = dat_f
  )

  mod <- metafor::rma.mv(
    yi = dat_f$lnVR, V = VCV,
    mods = ~ Outcome_type,
    random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
    test = "t", method = "REML", sparse = TRUE, data = dat_f
  )

  cf <- coef(summary(mod))

  tibble::tibble(
    rho              = r,
    anxiety_est      = cf["intrcpt", "estimate"],
    anxiety_ci.lb    = cf["intrcpt", "ci.lb"],
    anxiety_ci.ub    = cf["intrcpt", "ci.ub"],
    anxiety_pval     = cf["intrcpt", "pval"],
    depression_est   = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"],
    contrast_est     = cf["Outcome_typeDepression", "estimate"],
    contrast_ci.lb   = cf["Outcome_typeDepression", "ci.lb"],
    contrast_ci.ub   = cf["Outcome_typeDepression", "ci.ub"],
    contrast_tval    = cf["Outcome_typeDepression", "tval"],
    contrast_pval    = cf["Outcome_typeDepression", "pval"]
  )
})
```

```{r}
#| label: tbl_contrast_sensitivity
#| tbl-cap: "Depression-minus-anxiety lnVR contrast across assumed within-cohort correlations (rho)."

contrast_sensitivity %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: save_contrast_sensitivity_table
readr::write_csv(contrast_sensitivity, here("..","data","sensitivity_r_anxiety_depression_contrast.csv"))
```

## Interpretation

Across the entire plausible range tested (`rho = 0` to `rho = 0.90`):

-   The overall **lnRR** estimate stays negative and statistically significant (point estimate moves only from about -0.25 at `rho = 0` to about -0.23 at `rho = 0.90`; the 95% CI excludes zero at every value of `rho`). The substantive conclusion — that music exposure is associated with a reduction in anxiety-/depression-like behaviour — is **not sensitive** to the assumed within-cohort correlation.
-   The overall **lnVR** estimate stays small and its 95% CI includes zero at every value of `rho` tested. The conclusion that there is **no strong overall evidence for a change in behavioural variability** under music exposure is likewise **not sensitive** to `rho`.
-   The **depression-minus-anxiety lnVR contrast** remains negative and materially unchanged in magnitude (approximately -0.63 to -0.62 across the tested range) and remains statistically significant at every value of `rho`; the anxiety-specific pooled lnVR remains positive and nominally significant at `rho <= 0.75`, while the depression-specific pooled lnVR remains negative and non-significant throughout. The qualitative pattern reported in the manuscript (higher behavioural variability under music for anxiety-like outcomes than for depression-like outcomes) is **not driven by the specific choice of `rho = 0.5`**.

Increasing `rho` mechanically increases the assumed covariance between effect sizes sharing a cohort, which is absorbed mainly by a decrease in the `ES_ID`-level variance component and a modest narrowing of the standard errors of the pooled estimates; it does not change which cohorts or studies dominate the fit, and it does not alter the sign or statistical conclusion of the overall or contrast estimates reported here.

## Recomputing sampling variances across the assumed within-subject correlation ($r_i$)

`ri = 0.5` is used inside `calculate_lnRR()`, `calculate_lnVR()`, and `calculate_lnCVR()` (`effect_size_calculation.qmd`) as the assumed within-subject correlation between the treatment and control measurements of a *dependent* (paired / repeated-measures) comparison, via `metafor::escalc(measure = "ROMC"/"VRC"/"CVRC", ri = 0.5, ...)`. Only rows coded `Comparison_structure == "Dependent"` are affected. We re-confirm directly from the raw extraction database that this applies to exactly 40 of the dataset's 298 rows, and that all 40 come from a single study:

```{r}
#| label: verify_dependent_rows
#| message: false
#| warning: false

db_raw_check <- readr::read_csv(here("..","data","db251124.csv"), show_col_types = FALSE)

table(db_raw_check$Comparison_structure)

db_raw_check %>%
  filter(Comparison_structure == "Dependent") %>%
  count(Study_ID)
```

### Does $r_i$ change the point estimate or only the sampling variance?

`ROMC`, `VRC`, and `CVRC` are the "correlated / paired-groups" versions of the response-ratio, variability-ratio, and CV-ratio estimators. Algebraically, `ri` enters only the delta-method sampling-variance formula (through a covariance term that subtracts `2 * ri * sd1i * sd2i / (ni * ...)`) and never enters the point-estimate (`yi`) formula, which depends only on the means and/or SDs of the two groups. We confirm this empirically with a toy comparison, holding all other inputs fixed and varying only `ri`:

```{r}
#| label: verify_ri_affects_variance_only
#| message: false
#| warning: false

toy <- purrr::map_dfr(c(0.00, 0.25, 0.50, 0.75, 0.90), function(r) {
  purrr::map_dfr(c("ROMC", "VRC", "CVRC"), function(m) {
    e <- metafor::escalc(measure = m, ni = 12, m1i = 10, m2i = 8, sd1i = 2, sd2i = 1.5, ri = r)
    tibble::tibble(measure = m, ri = r, yi = as.numeric(e$yi), vi = as.numeric(e$vi))
  })
})

toy %>%
  kable(digits = 6, caption = "Point estimate (yi) is constant across ri; only the sampling variance (vi) changes.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))
```

As expected, `yi` is identical across `ri` for all three measures, while `vi` decreases monotonically as `ri` increases. This means that varying `ri` can only ever change the sampling variances (`lnRR_var`, `lnVR_var`, `lnCVR_var`) of the 40 dependent rows -- never their point estimates (`lnRR`, `lnVR`, `lnCVR`) -- and, by extension, can only change the pooled models through the weights it assigns to those 40 rows in the VCV/random-effects fit, not through the effect sizes themselves.

### Recomputing the 40 dependent-row sampling variances

We reproduce the same preprocessing that `effect_size_calculation.qmd` applies before calling `escalc()` -- SD imputation from SE where missing, and the `apply_continuity_within_group()` continuity correction (stratified by `Assay_type` x `Measurment_variable` x `Data_type`, applied to the *full* raw database so that the stratum-level epsilon constants match the ones used to build `data/db_effect_sizes.csv`) -- and then recompute `lnRR_var`, `lnVR_var`, and `lnCVR_var` for the 40 `Comparison_structure == "Dependent"` rows only, at `ri = 0.00, 0.25, 0.50, 0.75, 0.90`. Point estimates (`lnRR`, `lnVR`, `lnCVR`) and every value for the 258 independent-comparison rows are always taken unchanged from `data/db_effect_sizes.csv`, consistent with the result above that `ri` cannot affect them.

```{r}
#| label: continuity_helpers_ri
#| code-fold: TRUE

db_raw_ri <- readr::read_csv(here("..","data","db251124.csv"), show_col_types = FALSE) %>%
  mutate(
    across(c(C_n, Ex_n, C_mean, Ex_mean, C_SE, Ex_SE, C_SD, Ex_SD), as.numeric),
    Data_type = tolower(as.character(Data_type)),
    Comparison_structure = as.character(Comparison_structure),
    Higher_better = as.character(Higher_better)
  ) %>%
  mutate(
    C_SD  = ifelse(is.na(C_SD)  & !is.na(C_SE)  & !is.na(C_n),  C_SE  * sqrt(C_n),  C_SD),
    Ex_SD = ifelse(is.na(Ex_SD) & !is.na(Ex_SE) & !is.na(Ex_n), Ex_SE * sqrt(Ex_n), Ex_SD)
  ) %>%
  mutate(Cohort_ID = paste(Study_ID, Ex_ID, sep = "_"))

dep_ids_ri <- db_raw_ri %>% filter(Comparison_structure == "Dependent") %>% pull(ES_ID)

min_pos_ri <- function(x) {
  x <- x[is.finite(x) & !is.na(x) & x > 0]
  if (length(x) == 0) return(NA_real_)
  min(x)
}

apply_continuity_within_group_ri <- function(dt,
                                             strata = c("Assay_type", "Measurment_variable", "Data_type"),
                                             mult = 0.9) {
  dt <- dt %>%
    mutate(
      Data_type = tolower(as.character(Data_type)),
      across(c(C_mean, Ex_mean, C_SD, Ex_SD), ~as.numeric(as.character(.)))
    )
  mins <- dt %>%
    group_by(across(all_of(strata))) %>%
    summarise(
      min_C_mean_pos  = min_pos_ri(C_mean),
      min_Ex_mean_pos = min_pos_ri(Ex_mean),
      .groups = "drop"
    )
  g_C_mean  <- min_pos_ri(dt$C_mean)
  g_Ex_mean <- min_pos_ri(dt$Ex_mean)
  dt %>%
    left_join(mins, by = strata) %>%
    mutate(
      eps_C_mean  = mult * coalesce(min_C_mean_pos,  g_C_mean),
      eps_Ex_mean = mult * coalesce(min_Ex_mean_pos, g_Ex_mean),
      C_mean  = ifelse(!is.na(C_mean)  & C_mean == 0,  eps_C_mean,  C_mean),
      Ex_mean = ifelse(!is.na(Ex_mean) & Ex_mean == 0, eps_Ex_mean, Ex_mean),
      C_SD  = ifelse(!is.na(C_SD)  & C_SD == 0,  0.05, C_SD),
      Ex_SD = ifelse(!is.na(Ex_SD) & Ex_SD == 0, 0.05, Ex_SD)
    ) %>%
    select(-min_C_mean_pos, -min_Ex_mean_pos, -eps_C_mean, -eps_Ex_mean)
}
```

```{r}
#| label: recompute_dependent_var_ri
#| code-fold: TRUE
#| message: false
#| warning: false

recompute_dependent_var <- function(ri_val) {

  dt <- db_raw_ri %>% mutate(Data_type = tolower(as.character(Data_type)))

  dt_nonperc <- dt %>%
    filter(!Data_type %in% c("percentage")) %>%
    apply_continuity_within_group_ri()

  dt_perc <- dt %>%
    filter(Data_type %in% c("percentage")) %>%
    mutate(
      C_mean  = ifelse(C_mean  == 0,   0.5,  C_mean),
      Ex_mean = ifelse(Ex_mean == 0,   0.5,  Ex_mean),
      C_mean  = ifelse(C_mean  == 100, 99.5, C_mean),
      Ex_mean = ifelse(Ex_mean == 100, 99.5, Ex_mean),
      C_SD  = ifelse(C_SD  == 0, 0.5, C_SD),
      Ex_SD = ifelse(Ex_SD == 0, 0.5, Ex_SD)
    )

  asin_trans <- function(p) asin(sqrt(p))

  # lnRR_var: ROMC for continuous data types, manual paired-CV formula for percentage data
  d1 <- dt_nonperc %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff1 <- metafor::escalc(
    measure = "ROMC", ni = (d1$Ex_n + d1$C_n) / 2, m1i = d1$Ex_mean, m2i = d1$C_mean,
    sd1i = d1$Ex_SD, sd2i = d1$C_SD, ri = ri_val, var.names = c("y_", "v_")
  )
  lnRR_var_nonperc <- tibble(ES_ID = d1$ES_ID, lnRR_var_new = eff1$v_)

  d2 <- dt_perc %>% filter(ES_ID %in% dep_ids_ri)
  scale_flag <- pmax(d2$Ex_mean, d2$C_mean) > 1
  Ex_mean_r <- ifelse(scale_flag, d2$Ex_mean / 100, d2$Ex_mean)
  C_mean_r  <- ifelse(scale_flag, d2$C_mean  / 100, d2$C_mean)
  Ex_SD_r   <- ifelse(scale_flag, d2$Ex_SD   / 100, d2$Ex_SD)
  C_SD_r    <- ifelse(scale_flag, d2$C_SD    / 100, d2$C_SD)
  Ex_SD_trans <- sqrt(Ex_SD_r^2 / (4 * Ex_mean_r * (1 - Ex_mean_r)))
  C_SD_trans  <- sqrt(C_SD_r^2  / (4 * C_mean_r  * (1 - C_mean_r)))
  Ex_mean_trans <- asin_trans(Ex_mean_r)
  C_mean_trans  <- asin_trans(C_mean_r)
  V_Ex <- Ex_SD_trans^2 * (1 / (Ex_mean_trans^2 * d2$Ex_n))
  V_C  <- C_SD_trans^2  * (1 / (C_mean_trans^2  * d2$C_n))
  lnRR_var_perc <- tibble(ES_ID = d2$ES_ID, lnRR_var_new = V_Ex + V_C - 2 * ri_val * sqrt(V_Ex) * sqrt(V_C))

  lnRR_var_out <- bind_rows(lnRR_var_nonperc, lnRR_var_perc)

  # lnVR_var: VRC (non-percentage only)
  dt_vr <- db_raw_ri %>%
    mutate(Data_type = tolower(as.character(Data_type))) %>%
    filter(!Data_type %in% c("percentage")) %>%
    apply_continuity_within_group_ri()

  d3 <- dt_vr %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff3 <- metafor::escalc(
    measure = "VRC", ni = (d3$Ex_n + d3$C_n) / 2, sd1i = d3$Ex_SD, sd2i = d3$C_SD,
    ri = ri_val, var.names = c("y_", "v_")
  )
  lnVR_var_out <- tibble(ES_ID = d3$ES_ID, lnVR_var_new = eff3$v_)

  # lnCVR_var: CVRC (non-percentage only)
  d4 <- dt_vr %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff4 <- metafor::escalc(
    measure = "CVRC", ni = (d4$Ex_n + d4$C_n) / 2, m1i = d4$Ex_mean, m2i = d4$C_mean,
    sd1i = d4$Ex_SD, sd2i = d4$C_SD, ri = ri_val, var.names = c("y_", "v_")
  )
  lnCVR_var_out <- tibble(ES_ID = d4$ES_ID, lnCVR_var_new = eff4$v_)

  lnRR_var_out %>%
    full_join(lnVR_var_out, by = "ES_ID") %>%
    full_join(lnCVR_var_out, by = "ES_ID")
}

ri_values <- c(0.00, 0.25, 0.50, 0.75, 0.90)

# db_for_ri(): start from the analysis-ready effect-size table (db) and substitute only
# the *_var columns for the 40 dependent rows, at the given ri. Point estimates and all
# independent-row values are always taken unchanged from db.
db_for_ri <- function(ri_val) {
  sub <- recompute_dependent_var(ri_val) %>% mutate(ES_ID = as.character(ES_ID))
  db %>%
    mutate(ES_ID_chr = as.character(ES_ID)) %>%
    left_join(sub, by = c("ES_ID_chr" = "ES_ID")) %>%
    mutate(
      lnRR_var  = ifelse(!is.na(lnRR_var_new),  lnRR_var_new,  lnRR_var),
      lnVR_var  = ifelse(!is.na(lnVR_var_new),  lnVR_var_new,  lnVR_var),
      lnCVR_var = ifelse(!is.na(lnCVR_var_new), lnCVR_var_new, lnCVR_var)
    ) %>%
    select(-ES_ID_chr, -lnRR_var_new, -lnVR_var_new, -lnCVR_var_new)
}
```

**Validation check**: at `ri = 0.5` (the manuscript default), the recomputed sampling variances for the 40 dependent rows must reproduce `data/db_effect_sizes.csv`'s existing values exactly, since that file was itself built with `ri = 0.5`.

```{r}
#| label: validate_ri_var
#| message: false
#| warning: false

validation <- recompute_dependent_var(0.5) %>%
  mutate(ES_ID = as.character(ES_ID)) %>%
  left_join(
    db %>% filter(ES_ID %in% dep_ids_ri) %>%
      mutate(ES_ID = as.character(ES_ID)) %>%
      select(ES_ID, lnRR_var, lnVR_var, lnCVR_var),
    by = "ES_ID"
  )

tibble::tibble(
  column = c("lnRR_var", "lnVR_var", "lnCVR_var"),
  all.equal_to_db_effect_sizes = c(
    isTRUE(all.equal(validation$lnRR_var_new,  validation$lnRR_var,  tolerance = 1e-8)),
    isTRUE(all.equal(validation$lnVR_var_new[!is.na(validation$lnVR_var_new)],
                      validation$lnVR_var[!is.na(validation$lnVR_var_new)],  tolerance = 1e-8)),
    isTRUE(all.equal(validation$lnCVR_var_new[!is.na(validation$lnCVR_var_new)],
                      validation$lnCVR_var[!is.na(validation$lnCVR_var_new)], tolerance = 1e-8))
  )
) %>%
  kable(caption = "Validation: recomputed dependent-row sampling variances at ri = 0.5 vs. data/db_effect_sizes.csv.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))
```

The check passes: the recomputation pipeline exactly reproduces the existing `data/db_effect_sizes.csv` sampling variances at `ri = 0.5`, confirming the recomputed variances at other values of `ri` are directly comparable and attributable to the change in `ri` alone. (Point estimates are never recomputed here -- they are always carried over unchanged from `data/db_effect_sizes.csv`, which the result above shows is the correct and only defensible choice, since `ri` cannot affect them.)

### Refitting the overall models and the anxiety/depression contrast across $r_i$ (rho fixed at 0.5)

```{r}
#| label: fit_overall_ri
#| message: false
#| warning: false

overall_sensitivity_ri <- purrr::map_dfr(ri_values, function(r) {

  dset  <- db_for_ri(r)
  fit_rr <- fit_overall_mv(dset, "lnRR", "lnRR_var", 0.5)
  fit_vr <- fit_overall_mv(dset, "lnVR", "lnVR_var", 0.5)

  i2_rr <- orchaRd::i2_ml(fit_rr$model)
  i2_vr <- orchaRd::i2_ml(fit_vr$model)

  dplyr::bind_rows(
    tibble::tibble(
      metric        = "lnRR",
      ri            = r,
      k             = fit_rr$model$k,
      estimate      = as.numeric(coef(fit_rr$model)),
      se            = fit_rr$model$se,
      ci.lb         = fit_rr$model$ci.lb,
      ci.ub         = fit_rr$model$ci.ub,
      tval          = fit_rr$model$zval,
      pval          = fit_rr$model$pval,
      sigma2_Study  = fit_rr$model$sigma2[1],
      sigma2_ES     = fit_rr$model$sigma2[2],
      sigma2_Strain = fit_rr$model$sigma2[3],
      I2_Total      = i2_rr[["I2_Total"]],
      I2_Study_ID   = i2_rr[["I2_Study_ID"]],
      I2_ES_ID      = i2_rr[["I2_ES_ID"]],
      I2_Strain     = i2_rr[["I2_Strain"]]
    ),
    tibble::tibble(
      metric        = "lnVR",
      ri            = r,
      k             = fit_vr$model$k,
      estimate      = as.numeric(coef(fit_vr$model)),
      se            = fit_vr$model$se,
      ci.lb         = fit_vr$model$ci.lb,
      ci.ub         = fit_vr$model$ci.ub,
      tval          = fit_vr$model$zval,
      pval          = fit_vr$model$pval,
      sigma2_Study  = fit_vr$model$sigma2[1],
      sigma2_ES     = fit_vr$model$sigma2[2],
      sigma2_Strain = fit_vr$model$sigma2[3],
      I2_Total      = i2_vr[["I2_Total"]],
      I2_Study_ID   = i2_vr[["I2_Study_ID"]],
      I2_ES_ID      = i2_vr[["I2_ES_ID"]],
      I2_Strain     = i2_vr[["I2_Strain"]]
    )
  )
})
```

```{r}
#| label: tbl_overall_ri
#| tbl-cap: "Overall lnRR and lnVR pooled estimates across the assumed within-subject correlation (ri), rho fixed at 0.5."

overall_sensitivity_ri %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed")) %>%
  scroll_box(width = "100%")
```

```{r}
#| label: save_overall_ri
readr::write_csv(overall_sensitivity_ri, here("..","data","sensitivity_ri_overall_results.csv"))
```

```{r}
#| label: contrast_ri
#| message: false
#| warning: false

contrast_sensitivity_ri <- purrr::map_dfr(ri_values, function(r) {

  dset  <- db_for_ri(r)
  db_vr <- dset %>% filter(!is.na(lnVR), !is.na(lnVR_var))

  k_counts    <- db_vr %>% count(Outcome_type)
  keep_levels <- k_counts %>% filter(n >= 5) %>% pull(Outcome_type)

  dat_f <- db_vr %>%
    filter(Outcome_type %in% keep_levels) %>%
    mutate(Outcome_type = droplevels(Outcome_type))

  VCV <- metafor::vcalc(
    vi = dat_f$lnVR_var, cluster = dat_f$Cohort_ID, obs = dat_f$ES_ID,
    rho = 0.5, data = dat_f
  )

  mod <- metafor::rma.mv(
    yi = dat_f$lnVR, V = VCV,
    mods = ~ Outcome_type,
    random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
    test = "t", method = "REML", sparse = TRUE, data = dat_f
  )

  cf <- coef(summary(mod))

  tibble::tibble(
    ri               = r,
    anxiety_est      = cf["intrcpt", "estimate"],
    anxiety_ci.lb    = cf["intrcpt", "ci.lb"],
    anxiety_ci.ub    = cf["intrcpt", "ci.ub"],
    anxiety_pval     = cf["intrcpt", "pval"],
    depression_est   = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"],
    contrast_est     = cf["Outcome_typeDepression", "estimate"],
    contrast_ci.lb   = cf["Outcome_typeDepression", "ci.lb"],
    contrast_ci.ub   = cf["Outcome_typeDepression", "ci.ub"],
    contrast_tval    = cf["Outcome_typeDepression", "tval"],
    contrast_pval    = cf["Outcome_typeDepression", "pval"]
  )
})
```

```{r}
#| label: tbl_contrast_ri
#| tbl-cap: "Depression-minus-anxiety lnVR contrast across the assumed within-subject correlation (ri), rho fixed at 0.5."

contrast_sensitivity_ri %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: save_contrast_ri
readr::write_csv(contrast_sensitivity_ri, here("..","data","sensitivity_ri_anxiety_depression_contrast.csv"))
```

```{r}
#| label: fig_overall_ri
#| fig-width: 9
#| fig-height: 5

p_sensitivity_ri <- overall_sensitivity_ri %>%
  mutate(metric = factor(metric, levels = c("lnRR", "lnVR"))) %>%
  ggplot(aes(x = factor(ri), y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci.lb, ymax = ci.ub), size = 0.6) +
  facet_wrap(~ metric, scales = "free_y") +
  labs(
    x = expression(paste("Assumed within-subject correlation, ", r[i])),
    y = "Pooled estimate (95% CI)",
    title = "Overall lnRR and lnVR estimates across assumed values of ri (rho = 0.5)"
  ) +
  theme_bw(base_size = 13)

p_sensitivity_ri
```

```{r}
#| label: save-figure_sensitivity_ri
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_ri_overall.pdf"),
       plot = p_sensitivity_ri,
       dpi = 300, device = cairo_pdf, width = 9, height = 5)

ggsave(filename = here("..","Plots", "sensitivity_ri_overall.jpg"),
       plot = p_sensitivity_ri,
       width = 9, height = 5, dpi = 300)
```

## Joint sensitivity to $r_i$ and $\rho$

Because `ri` and `rho` are conceptually independent (one governs the sampling variance of 40 within-subject comparisons; the other governs the assumed covariance between different effect sizes sharing a cohort), we also cross them in a full 5 x 5 grid (`ri = 0.00, 0.25, 0.50, 0.75, 0.90` by `rho = 0.00, 0.25, 0.50, 0.75, 0.90`, 25 combinations) and refit the overall `lnRR`, overall `lnVR`, and the depression-minus-anxiety `lnVR` contrast at each combination (same model specification throughout).

```{r}
#| label: grid_ri_rho
#| message: false
#| warning: false

grid_ri_rho_results <- purrr::map_dfr(ri_values, function(r_i) {

  dset <- db_for_ri(r_i)

  purrr::map_dfr(rho_values, function(r_rho) {

    fit_rr <- fit_overall_mv(dset, "lnRR", "lnRR_var", r_rho)
    fit_vr <- fit_overall_mv(dset, "lnVR", "lnVR_var", r_rho)

    db_vr_g <- dset %>% filter(!is.na(lnVR), !is.na(lnVR_var))
    keep_levels_g <- db_vr_g %>% count(Outcome_type) %>% filter(n >= 5) %>% pull(Outcome_type)
    dat_fg <- db_vr_g %>% filter(Outcome_type %in% keep_levels_g) %>% mutate(Outcome_type = droplevels(Outcome_type))
    VCV_g <- metafor::vcalc(vi = dat_fg$lnVR_var, cluster = dat_fg$Cohort_ID, obs = dat_fg$ES_ID, rho = r_rho, data = dat_fg)
    mod_ct <- metafor::rma.mv(
      yi = dat_fg$lnVR, V = VCV_g, mods = ~ Outcome_type,
      random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
      test = "t", method = "REML", sparse = TRUE, data = dat_fg
    )
    cf <- coef(summary(mod_ct))

    dplyr::bind_rows(
      tibble::tibble(model = "lnRR_overall", ri = r_i, rho = r_rho, k = fit_rr$model$k,
                      estimate = as.numeric(coef(fit_rr$model)), se = fit_rr$model$se,
                      ci.lb = fit_rr$model$ci.lb, ci.ub = fit_rr$model$ci.ub,
                      tval = fit_rr$model$zval, pval = fit_rr$model$pval,
                      anxiety_est = NA_real_, depression_est = NA_real_),
      tibble::tibble(model = "lnVR_overall", ri = r_i, rho = r_rho, k = fit_vr$model$k,
                      estimate = as.numeric(coef(fit_vr$model)), se = fit_vr$model$se,
                      ci.lb = fit_vr$model$ci.lb, ci.ub = fit_vr$model$ci.ub,
                      tval = fit_vr$model$zval, pval = fit_vr$model$pval,
                      anxiety_est = NA_real_, depression_est = NA_real_),
      tibble::tibble(model = "lnVR_contrast_depression_minus_anxiety", ri = r_i, rho = r_rho, k = mod_ct$k,
                      estimate = cf["Outcome_typeDepression", "estimate"], se = cf["Outcome_typeDepression", "se"],
                      ci.lb = cf["Outcome_typeDepression", "ci.lb"], ci.ub = cf["Outcome_typeDepression", "ci.ub"],
                      tval = cf["Outcome_typeDepression", "tval"], pval = cf["Outcome_typeDepression", "pval"],
                      anxiety_est = cf["intrcpt", "estimate"],
                      depression_est = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"])
    )
  })
})

readr::write_csv(grid_ri_rho_results, here("..","data","sensitivity_ri_rho_grid_results.csv"))
```

```{r}
#| label: tbl_grid_summary
#| tbl-cap: "Summary of the 25-combination ri x rho grid: range of pooled estimates and whether the 95% CI ever crosses zero (lnRR/contrast) or ever excludes zero (lnVR)."

grid_summary <- grid_ri_rho_results %>%
  group_by(model) %>%
  summarise(
    estimate_min   = round(min(estimate), 4),
    estimate_max   = round(max(estimate), 4),
    pval_min       = signif(min(pval), 3),
    pval_max       = signif(max(pval), 3),
    ci_crosses_zero    = any(ci.lb <= 0 & ci.ub >= 0),
    ci_excludes_zero_always = all(ci.lb > 0 | ci.ub < 0),
    sign_ever_flips = any(sign(estimate) != sign(estimate[1])),
    .groups = "drop"
  )

grid_summary %>%
  kable(caption = "Grid summary across all 25 ri x rho combinations.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: fig_grid_heatmap
#| fig-width: 10
#| fig-height: 4

model_labels <- c(
  lnRR_overall = "Overall lnRR",
  lnVR_overall = "Overall lnVR",
  lnVR_contrast_depression_minus_anxiety = "Depression - anxiety lnVR contrast"
)

p_grid <- grid_ri_rho_results %>%
  mutate(model_label = factor(model_labels[model], levels = model_labels)) %>%
  ggplot(aes(x = factor(ri), y = factor(rho), fill = estimate)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = round(estimate, 2)), size = 3) +
  facet_wrap(~ model_label, scales = "free") +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b", midpoint = 0) +
  labs(
    x = expression(paste("Assumed within-subject correlation, ", r[i])),
    y = expression(paste("Assumed within-cohort correlation, ", rho)),
    fill = "Estimate",
    title = "Pooled estimates across the joint ri x rho grid (25 combinations)"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right")

p_grid
```

```{r}
#| label: save_fig_grid
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_ri_rho_grid.pdf"),
       plot = p_grid,
       dpi = 300, device = cairo_pdf, width = 10, height = 4)

ggsave(filename = here("..","Plots", "sensitivity_ri_rho_grid.jpg"),
       plot = p_grid,
       width = 10, height = 4, dpi = 300)
```

## Interpretation of the $r_i$ and joint $r_i \times \rho$ sensitivity

Varying `ri` from 0 to 0.90 (rho fixed at the manuscript default of 0.5) changes only the sampling variances of the 40 dependent rows from a single study (`Sampaio_2017_PSYNEURO`), and the resulting shifts in the pooled estimates are small:

-   The overall **lnRR** estimate moves only from about -0.237 (`ri = 0`) to about -0.234 (`ri = 0.90`), staying negative and statistically significant throughout (95% CI excludes zero at every value of `ri`; p ranges from about 0.0006 to 0.0007).
-   The overall **lnVR** estimate moves only from about 0.132 (`ri = 0`) to about 0.127 (`ri = 0.90`); its 95% CI includes zero at every value of `ri`, so the conclusion of no strong overall evidence for a change in behavioural variability is unaffected.
-   The **depression-minus-anxiety lnVR contrast** moves only from about -0.626 (`ri = 0`) to about -0.616 (`ri = 0.90`) and remains statistically significant throughout (p from about 0.005 to 0.006); the anxiety-specific pooled lnVR stays positive and nominally significant, and the depression-specific pooled lnVR stays negative and non-significant, across the full range of `ri`.

This is expected given the algebraic/empirical result above: `ri` affects only the *weights* the 40 dependent rows receive in the multilevel model, not their point estimates, and those 40 rows are a small fraction (13%) of the 298-row database, so even a large swing in `ri` has limited leverage on the pooled estimates.

Crossing `ri` with `rho` in the full 25-combination grid does not change this picture:

-   **Overall lnRR** ranges from about -0.250 to -0.226 across all 25 combinations, and its 95% CI excludes zero in every combination (p from about 0.0005 to 0.0011): the reduction in anxiety-/depression-like behaviour under music exposure is robust to the joint choice of `ri` and `rho`.
-   **Overall lnVR** ranges from about 0.116 to 0.141 across all 25 combinations, and its 95% CI includes zero in every combination: there remains no strong overall evidence for a change in behavioural variability under music exposure, regardless of `ri` and `rho`.
-   The **depression-minus-anxiety lnVR contrast** ranges from about -0.639 to -0.616 across all 25 combinations, never changes sign, and never loses statistical significance (p stays below 0.006 throughout, well under the 0.05 threshold); the anxiety-specific pooled lnVR ranges from about 0.29 to 0.31 and the depression-specific pooled lnVR from about -0.32 to -0.31 across the grid.

In short, none of the three substantive conclusions reported in the manuscript -- a significant overall reduction in anxiety-/depression-like behaviour (lnRR), no strong overall change in behavioural variability (lnVR), and a significantly more positive lnVR response for anxiety-like than for depression-like outcomes -- depends on the specific default choice of `ri = 0.5` or `rho = 0.5`, individually or jointly.

::: callout-note
```{r}
sessionInfo()
```
:::
