---
title: "Sensitivity analysis to the assumed within-study correlation"
---

The overall-effect and moderator models (see `overall_effect.qmd` and `uni_moderators.qmd`) need an assumption about how correlated the sampling errors of different effect sizes are, so that effect sizes which are not really independent of one another are not treated as if they were. Two separate correlation assumptions are used in the pipeline, and both happen to be set to the same round number, 0.5, by convention -- but they answer different questions:

1.  **Correlation between different outcomes from the same cohort** (`vcalc(rho = 0.5)`). Some cohorts of animals contribute more than one effect size to the dataset (57 of them do) -- for example, two different behavioural tests run on the same group of animals. Effect sizes coming from the same cohort are not statistically independent, and `rho` is our assumption about how strongly they are related. This assumption feeds into the off-diagonal entries of the variance-covariance (VCV) matrix used by every `rma.mv()` model in the analysis.
2.  **Correlation between the two measurements inside a single paired comparison** (`escalc(ri = 0.5)`). Some comparisons in the dataset are paired / repeated-measures designs, where the treatment and control measurements come from the same animals rather than from separate groups. There are 40 such effect sizes in the database, all from a single study (`Sampaio_2017_PSYNEURO`, a within-subject design measured at four life stages), coded `Comparison_structure == "Dependent"`. `ri` is our assumption about how correlated those paired measurements are, and it only affects the sampling-variance formula for those 40 rows.

Because both assumptions are round-number defaults rather than values estimated from the data, this appendix checks whether the models' results depend on the specific choice of 0.5. We first vary `rho` for the overall `lnRR`/`lnVR` models, since that is the parameter that structures the VCV matrix referenced in the manuscript. Further down, in "Recomputing sampling variances across the assumed within-subject correlation", we vary `ri` for the 40 affected rows, recomputing their sampling variances directly from the raw extraction database (`data/db251124.csv`). Finally, we vary both parameters together in a joint `ri` x `rho` grid.

```{r}
#| label: setup
#| include: false
pacman::p_load(
               DT,
               dtplyr,
               here,
               knitr,
               tidyverse,
               patchwork,
               metafor,
               orchaRd,
               kableExtra
               )
```

```{r}
#| label: load_effectsize_db
#| message: false
#| warning: false

db <- readr::read_csv(here("..","data","db_effect_sizes.csv")) %>%
  mutate(across(where(is.character), as.factor)) %>%
  mutate(across(c(lnRR, lnRR_var, lnVR, lnVR_var, lnCVR, lnCVR_var), as.numeric))
```

## Refitting the overall models across a range of assumed correlations

To see whether the choice of `rho` matters, we refit the exact same intercept-only multilevel models used in `overall_effect.qmd` (random intercepts for `Study_ID`, `ES_ID`, and `Strain`; REML; `test = "t"`), changing only the `rho` value passed to `vcalc()` when building the VCV matrix. Everything else about the model is held fixed. We try five values of `rho`, spanning from assuming no correlation at all (`rho = 0`) to assuming a strong correlation (`rho = 0.90`): `0.00, 0.25, 0.50, 0.75, 0.90`.

```{r}
#| label: fit_overall_mv_rho
#| message: false
#| warning: false

rho_values <- c(0.00, 0.25, 0.50, 0.75, 0.90)

fit_overall_mv <- function(dat, yi, vi, rho) {

  dat2 <- dat %>%
    filter(!is.na(.data[[yi]]), !is.na(.data[[vi]]))

  VCV <- metafor::vcalc(
    vi = dat2[[vi]],
    cluster = dat2$Cohort_ID,
    obs = dat2$ES_ID,
    rho = rho,
    data = dat2
  )

  mod <- metafor::rma.mv(
    yi = dat2[[yi]],
    V  = VCV,
    random = list(
      ~1 | Study_ID,
      ~1 | ES_ID,
      ~1 | Strain
    ),
    test = "t",
    method = "REML",
    sparse = TRUE,
    data = dat2
  )

  list(data = dat2, VCV = VCV, model = mod)
}

overall_sensitivity <- purrr::map_dfr(rho_values, function(r) {

  fit_rr <- fit_overall_mv(db, "lnRR", "lnRR_var", r)
  fit_vr <- fit_overall_mv(db, "lnVR", "lnVR_var", r)

  i2_rr <- orchaRd::i2_ml(fit_rr$model)
  i2_vr <- orchaRd::i2_ml(fit_vr$model)

  dplyr::bind_rows(
    tibble::tibble(
      metric        = "lnRR",
      rho           = r,
      k             = fit_rr$model$k,
      estimate      = as.numeric(coef(fit_rr$model)),
      se            = fit_rr$model$se,
      ci.lb         = fit_rr$model$ci.lb,
      ci.ub         = fit_rr$model$ci.ub,
      tval          = fit_rr$model$zval,
      pval          = fit_rr$model$pval,
      sigma2_Study  = fit_rr$model$sigma2[1],
      sigma2_ES     = fit_rr$model$sigma2[2],
      sigma2_Strain = fit_rr$model$sigma2[3],
      I2_Total      = i2_rr[["I2_Total"]],
      I2_Study_ID   = i2_rr[["I2_Study_ID"]],
      I2_ES_ID      = i2_rr[["I2_ES_ID"]],
      I2_Strain     = i2_rr[["I2_Strain"]]
    ),
    tibble::tibble(
      metric        = "lnVR",
      rho           = r,
      k             = fit_vr$model$k,
      estimate      = as.numeric(coef(fit_vr$model)),
      se            = fit_vr$model$se,
      ci.lb         = fit_vr$model$ci.lb,
      ci.ub         = fit_vr$model$ci.ub,
      tval          = fit_vr$model$zval,
      pval          = fit_vr$model$pval,
      sigma2_Study  = fit_vr$model$sigma2[1],
      sigma2_ES     = fit_vr$model$sigma2[2],
      sigma2_Strain = fit_vr$model$sigma2[3],
      I2_Total      = i2_vr[["I2_Total"]],
      I2_Study_ID   = i2_vr[["I2_Study_ID"]],
      I2_ES_ID      = i2_vr[["I2_ES_ID"]],
      I2_Strain     = i2_vr[["I2_Strain"]]
    )
  )
})
```

```{r}
#| label: tbl_overall_sensitivity
#| tbl-cap: "Overall lnRR and lnVR pooled estimates across assumed within-cohort correlations (rho)."

overall_sensitivity %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed")) %>%
  scroll_box(width = "100%")
```

```{r}
#| label: save_overall_sensitivity_table
readr::write_csv(overall_sensitivity, here("..","data","sensitivity_r_overall_results.csv"))
```

## Figure: overall estimate and 95% CI across rho

```{r}
#| label: fig_overall_sensitivity
#| fig-width: 9
#| fig-height: 5

p_sensitivity <- overall_sensitivity %>%
  mutate(metric = factor(metric, levels = c("lnRR", "lnVR"))) %>%
  ggplot(aes(x = factor(rho), y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci.lb, ymax = ci.ub), size = 0.6) +
  facet_wrap(~ metric, scales = "free_y") +
  labs(
    x = expression(paste("Assumed within-cohort correlation, ", rho)),
    y = "Pooled estimate (95% CI)",
    title = "Overall lnRR and lnVR estimates across assumed values of the VCV correlation"
  ) +
  theme_bw(base_size = 13)

p_sensitivity
```

```{r}
#| label: save-figure_sensitivity
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_r_overall.pdf"),
       plot = p_sensitivity,
       dpi = 300, device = cairo_pdf, width = 9, height = 5)

ggsave(filename = here("..","Plots", "sensitivity_r_overall.jpg"),
       plot = p_sensitivity,
       width = 9, height = 5, dpi = 300)
```

## Anxiety-versus-depression lnVR contrast across rho

The manuscript also reports a depression-minus-anxiety contrast for `lnVR` (see `uni_moderators.qmd`, "Outcome type" section). Here we check whether that contrast -- and the two subgroup pooled estimates it is built from -- changes as `rho` changes. We apply the same rule used in `uni_moderators.qmd` to drop sparse levels (the `both` level of `Outcome_type`, which only has `k = 3`, is dropped).

```{r}
#| label: contrast_sensitivity
#| message: false
#| warning: false

db_vr <- db %>% filter(!is.na(lnVR), !is.na(lnVR_var))

k_counts    <- db_vr %>% count(Outcome_type)
keep_levels <- k_counts %>% filter(n >= 5) %>% pull(Outcome_type)

dat_f <- db_vr %>%
  filter(Outcome_type %in% keep_levels) %>%
  mutate(Outcome_type = droplevels(Outcome_type))

contrast_sensitivity <- purrr::map_dfr(rho_values, function(r) {

  VCV <- metafor::vcalc(
    vi = dat_f$lnVR_var, cluster = dat_f$Cohort_ID, obs = dat_f$ES_ID,
    rho = r, data = dat_f
  )

  mod <- metafor::rma.mv(
    yi = dat_f$lnVR, V = VCV,
    mods = ~ Outcome_type,
    random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
    test = "t", method = "REML", sparse = TRUE, data = dat_f
  )

  cf <- coef(summary(mod))

  tibble::tibble(
    rho              = r,
    anxiety_est      = cf["intrcpt", "estimate"],
    anxiety_ci.lb    = cf["intrcpt", "ci.lb"],
    anxiety_ci.ub    = cf["intrcpt", "ci.ub"],
    anxiety_pval     = cf["intrcpt", "pval"],
    depression_est   = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"],
    contrast_est     = cf["Outcome_typeDepression", "estimate"],
    contrast_ci.lb   = cf["Outcome_typeDepression", "ci.lb"],
    contrast_ci.ub   = cf["Outcome_typeDepression", "ci.ub"],
    contrast_tval    = cf["Outcome_typeDepression", "tval"],
    contrast_pval    = cf["Outcome_typeDepression", "pval"]
  )
})
```

```{r}
#| label: tbl_contrast_sensitivity
#| tbl-cap: "Depression-minus-anxiety lnVR contrast across assumed within-cohort correlations (rho)."

contrast_sensitivity %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: save_contrast_sensitivity_table
readr::write_csv(contrast_sensitivity, here("..","data","sensitivity_r_anxiety_depression_contrast.csv"))
```

## Recomputing sampling variances across the assumed within-subject correlation ($r_i$)

`ri = 0.5` is the assumed within-subject correlation used inside `calculate_lnRR()`, `calculate_lnVR()`, and `calculate_lnCVR()` (`effect_size_calculation.qmd`) for the 40 paired / repeated-measures comparisons, via `metafor::escalc(measure = "ROMC"/"VRC"/"CVRC", ri = 0.5, ...)`. Only rows coded `Comparison_structure == "Dependent"` are affected. Before varying `ri`, we first re-confirm directly from the raw extraction database that this applies to exactly 40 of the dataset's 298 rows, and that all 40 come from a single study:

```{r}
#| label: verify_dependent_rows
#| message: false
#| warning: false

db_raw_check <- readr::read_csv(here("..","data","db251124.csv"), show_col_types = FALSE)

table(db_raw_check$Comparison_structure)

db_raw_check %>%
  filter(Comparison_structure == "Dependent") %>%
  count(Study_ID)
```

### Does $r_i$ change the point estimate or only the sampling variance?

`ROMC`, `VRC`, and `CVRC` are the paired-groups versions of the response-ratio, variability-ratio, and CV-ratio formulas. `ri` only shows up in the sampling-variance formula (through a term that subtracts a covariance component); it never shows up in the formula for the point estimate itself, which depends only on the means and/or SDs of the two groups. We check this directly with a toy example, holding every other input fixed and changing only `ri`:

```{r}
#| label: verify_ri_affects_variance_only
#| message: false
#| warning: false

toy <- purrr::map_dfr(c(0.00, 0.25, 0.50, 0.75, 0.90), function(r) {
  purrr::map_dfr(c("ROMC", "VRC", "CVRC"), function(m) {
    e <- metafor::escalc(measure = m, ni = 12, m1i = 10, m2i = 8, sd1i = 2, sd2i = 1.5, ri = r)
    tibble::tibble(measure = m, ri = r, yi = as.numeric(e$yi), vi = as.numeric(e$vi))
  })
})

toy %>%
  kable(digits = 6, caption = "Point estimate (yi) is constant across ri; only the sampling variance (vi) changes.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))
```

As expected, `yi` stays identical across `ri` for all three measures, while `vi` gets smaller as `ri` increases. So changing `ri` can only ever change the sampling variances (`lnRR_var`, `lnVR_var`, `lnCVR_var`) of the 40 dependent rows -- never their point estimates (`lnRR`, `lnVR`, `lnCVR`) -- and it can only affect the pooled models through the weight those 40 rows get in the fit, not through the effect sizes themselves.

### Recomputing the 40 dependent-row sampling variances

To vary `ri`, we reproduce the same preprocessing steps that `effect_size_calculation.qmd` applies before calling `escalc()` -- imputing SD from SE where it is missing, and applying the same continuity correction (`apply_continuity_within_group()`, stratified by `Assay_type` x `Measurment_variable` x `Data_type`, and run on the *full* raw database so the stratum-level epsilon constants match the ones used to build `data/db_effect_sizes.csv`) -- and then recompute `lnRR_var`, `lnVR_var`, and `lnCVR_var` for the 40 `Comparison_structure == "Dependent"` rows only, at `ri = 0.00, 0.25, 0.50, 0.75, 0.90`. Point estimates (`lnRR`, `lnVR`, `lnCVR`) and every value for the 258 independent-comparison rows are always taken unchanged from `data/db_effect_sizes.csv`, since the result above shows `ri` cannot affect them.

```{r}
#| label: continuity_helpers_ri
#| code-fold: TRUE

db_raw_ri <- readr::read_csv(here("..","data","db251124.csv"), show_col_types = FALSE) %>%
  mutate(
    across(c(C_n, Ex_n, C_mean, Ex_mean, C_SE, Ex_SE, C_SD, Ex_SD), as.numeric),
    Data_type = tolower(as.character(Data_type)),
    Comparison_structure = as.character(Comparison_structure),
    Higher_better = as.character(Higher_better)
  ) %>%
  mutate(
    C_SD  = ifelse(is.na(C_SD)  & !is.na(C_SE)  & !is.na(C_n),  C_SE  * sqrt(C_n),  C_SD),
    Ex_SD = ifelse(is.na(Ex_SD) & !is.na(Ex_SE) & !is.na(Ex_n), Ex_SE * sqrt(Ex_n), Ex_SD)
  ) %>%
  mutate(Cohort_ID = paste(Study_ID, Ex_ID, sep = "_"))

dep_ids_ri <- db_raw_ri %>% filter(Comparison_structure == "Dependent") %>% pull(ES_ID)

min_pos_ri <- function(x) {
  x <- x[is.finite(x) & !is.na(x) & x > 0]
  if (length(x) == 0) return(NA_real_)
  min(x)
}

apply_continuity_within_group_ri <- function(dt,
                                             strata = c("Assay_type", "Measurment_variable", "Data_type"),
                                             mult = 0.9) {
  dt <- dt %>%
    mutate(
      Data_type = tolower(as.character(Data_type)),
      across(c(C_mean, Ex_mean, C_SD, Ex_SD), ~as.numeric(as.character(.)))
    )
  mins <- dt %>%
    group_by(across(all_of(strata))) %>%
    summarise(
      min_C_mean_pos  = min_pos_ri(C_mean),
      min_Ex_mean_pos = min_pos_ri(Ex_mean),
      .groups = "drop"
    )
  g_C_mean  <- min_pos_ri(dt$C_mean)
  g_Ex_mean <- min_pos_ri(dt$Ex_mean)
  dt %>%
    left_join(mins, by = strata) %>%
    mutate(
      eps_C_mean  = mult * coalesce(min_C_mean_pos,  g_C_mean),
      eps_Ex_mean = mult * coalesce(min_Ex_mean_pos, g_Ex_mean),
      C_mean  = ifelse(!is.na(C_mean)  & C_mean == 0,  eps_C_mean,  C_mean),
      Ex_mean = ifelse(!is.na(Ex_mean) & Ex_mean == 0, eps_Ex_mean, Ex_mean),
      C_SD  = ifelse(!is.na(C_SD)  & C_SD == 0,  0.05, C_SD),
      Ex_SD = ifelse(!is.na(Ex_SD) & Ex_SD == 0, 0.05, Ex_SD)
    ) %>%
    select(-min_C_mean_pos, -min_Ex_mean_pos, -eps_C_mean, -eps_Ex_mean)
}
```

```{r}
#| label: recompute_dependent_var_ri
#| code-fold: TRUE
#| message: false
#| warning: false

recompute_dependent_var <- function(ri_val) {

  dt <- db_raw_ri %>% mutate(Data_type = tolower(as.character(Data_type)))

  dt_nonperc <- dt %>%
    filter(!Data_type %in% c("percentage")) %>%
    apply_continuity_within_group_ri()

  dt_perc <- dt %>%
    filter(Data_type %in% c("percentage")) %>%
    mutate(
      C_mean  = ifelse(C_mean  == 0,   0.5,  C_mean),
      Ex_mean = ifelse(Ex_mean == 0,   0.5,  Ex_mean),
      C_mean  = ifelse(C_mean  == 100, 99.5, C_mean),
      Ex_mean = ifelse(Ex_mean == 100, 99.5, Ex_mean),
      C_SD  = ifelse(C_SD  == 0, 0.5, C_SD),
      Ex_SD = ifelse(Ex_SD == 0, 0.5, Ex_SD)
    )

  asin_trans <- function(p) asin(sqrt(p))

  # lnRR_var: ROMC for continuous data types, manual paired-CV formula for percentage data
  d1 <- dt_nonperc %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff1 <- metafor::escalc(
    measure = "ROMC", ni = (d1$Ex_n + d1$C_n) / 2, m1i = d1$Ex_mean, m2i = d1$C_mean,
    sd1i = d1$Ex_SD, sd2i = d1$C_SD, ri = ri_val, var.names = c("y_", "v_")
  )
  lnRR_var_nonperc <- tibble(ES_ID = d1$ES_ID, lnRR_var_new = eff1$v_)

  d2 <- dt_perc %>% filter(ES_ID %in% dep_ids_ri)
  scale_flag <- pmax(d2$Ex_mean, d2$C_mean) > 1
  Ex_mean_r <- ifelse(scale_flag, d2$Ex_mean / 100, d2$Ex_mean)
  C_mean_r  <- ifelse(scale_flag, d2$C_mean  / 100, d2$C_mean)
  Ex_SD_r   <- ifelse(scale_flag, d2$Ex_SD   / 100, d2$Ex_SD)
  C_SD_r    <- ifelse(scale_flag, d2$C_SD    / 100, d2$C_SD)
  Ex_SD_trans <- sqrt(Ex_SD_r^2 / (4 * Ex_mean_r * (1 - Ex_mean_r)))
  C_SD_trans  <- sqrt(C_SD_r^2  / (4 * C_mean_r  * (1 - C_mean_r)))
  Ex_mean_trans <- asin_trans(Ex_mean_r)
  C_mean_trans  <- asin_trans(C_mean_r)
  V_Ex <- Ex_SD_trans^2 * (1 / (Ex_mean_trans^2 * d2$Ex_n))
  V_C  <- C_SD_trans^2  * (1 / (C_mean_trans^2  * d2$C_n))
  lnRR_var_perc <- tibble(ES_ID = d2$ES_ID, lnRR_var_new = V_Ex + V_C - 2 * ri_val * sqrt(V_Ex) * sqrt(V_C))

  lnRR_var_out <- bind_rows(lnRR_var_nonperc, lnRR_var_perc)

  # lnVR_var: VRC (non-percentage only)
  dt_vr <- db_raw_ri %>%
    mutate(Data_type = tolower(as.character(Data_type))) %>%
    filter(!Data_type %in% c("percentage")) %>%
    apply_continuity_within_group_ri()

  d3 <- dt_vr %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff3 <- metafor::escalc(
    measure = "VRC", ni = (d3$Ex_n + d3$C_n) / 2, sd1i = d3$Ex_SD, sd2i = d3$C_SD,
    ri = ri_val, var.names = c("y_", "v_")
  )
  lnVR_var_out <- tibble(ES_ID = d3$ES_ID, lnVR_var_new = eff3$v_)

  # lnCVR_var: CVRC (non-percentage only)
  d4 <- dt_vr %>%
    filter(ES_ID %in% dep_ids_ri, Data_type %in% c("time", "distance", "count", "latency", "index"))
  eff4 <- metafor::escalc(
    measure = "CVRC", ni = (d4$Ex_n + d4$C_n) / 2, m1i = d4$Ex_mean, m2i = d4$C_mean,
    sd1i = d4$Ex_SD, sd2i = d4$C_SD, ri = ri_val, var.names = c("y_", "v_")
  )
  lnCVR_var_out <- tibble(ES_ID = d4$ES_ID, lnCVR_var_new = eff4$v_)

  lnRR_var_out %>%
    full_join(lnVR_var_out, by = "ES_ID") %>%
    full_join(lnCVR_var_out, by = "ES_ID")
}

ri_values <- c(0.00, 0.25, 0.50, 0.75, 0.90)

# db_for_ri(): start from the analysis-ready effect-size table (db) and substitute only
# the *_var columns for the 40 dependent rows, at the given ri. Point estimates and all
# independent-row values are always taken unchanged from db.
db_for_ri <- function(ri_val) {
  sub <- recompute_dependent_var(ri_val) %>% mutate(ES_ID = as.character(ES_ID))
  db %>%
    mutate(ES_ID_chr = as.character(ES_ID)) %>%
    left_join(sub, by = c("ES_ID_chr" = "ES_ID")) %>%
    mutate(
      lnRR_var  = ifelse(!is.na(lnRR_var_new),  lnRR_var_new,  lnRR_var),
      lnVR_var  = ifelse(!is.na(lnVR_var_new),  lnVR_var_new,  lnVR_var),
      lnCVR_var = ifelse(!is.na(lnCVR_var_new), lnCVR_var_new, lnCVR_var)
    ) %>%
    select(-ES_ID_chr, -lnRR_var_new, -lnVR_var_new, -lnCVR_var_new)
}
```

**Validation check**: at `ri = 0.5`, the manuscript's default, this recomputation should exactly reproduce the sampling variances already stored in `data/db_effect_sizes.csv` for the 40 dependent rows, since that file was itself built with `ri = 0.5`.

```{r}
#| label: validate_ri_var
#| message: false
#| warning: false

validation <- recompute_dependent_var(0.5) %>%
  mutate(ES_ID = as.character(ES_ID)) %>%
  left_join(
    db %>% filter(ES_ID %in% dep_ids_ri) %>%
      mutate(ES_ID = as.character(ES_ID)) %>%
      select(ES_ID, lnRR_var, lnVR_var, lnCVR_var),
    by = "ES_ID"
  )

tibble::tibble(
  column = c("lnRR_var", "lnVR_var", "lnCVR_var"),
  all.equal_to_db_effect_sizes = c(
    isTRUE(all.equal(validation$lnRR_var_new,  validation$lnRR_var,  tolerance = 1e-8)),
    isTRUE(all.equal(validation$lnVR_var_new[!is.na(validation$lnVR_var_new)],
                      validation$lnVR_var[!is.na(validation$lnVR_var_new)],  tolerance = 1e-8)),
    isTRUE(all.equal(validation$lnCVR_var_new[!is.na(validation$lnCVR_var_new)],
                      validation$lnCVR_var[!is.na(validation$lnCVR_var_new)], tolerance = 1e-8))
  )
) %>%
  kable(caption = "Validation: recomputed dependent-row sampling variances at ri = 0.5 vs. data/db_effect_sizes.csv.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed"))
```

The check passes: the recomputation reproduces the existing `data/db_effect_sizes.csv` sampling variances exactly at `ri = 0.5`. This confirms the pipeline is doing what it is meant to, so the variances recomputed at other values of `ri` are directly comparable, and any differences from the current results are attributable to the change in `ri` alone. Point estimates are never recomputed here -- they are always carried over unchanged from `data/db_effect_sizes.csv`, consistent with `ri` not affecting them.

### Refitting the overall models and the anxiety/depression contrast across $r_i$ (rho fixed at 0.5)

```{r}
#| label: fit_overall_ri
#| message: false
#| warning: false

overall_sensitivity_ri <- purrr::map_dfr(ri_values, function(r) {

  dset  <- db_for_ri(r)
  fit_rr <- fit_overall_mv(dset, "lnRR", "lnRR_var", 0.5)
  fit_vr <- fit_overall_mv(dset, "lnVR", "lnVR_var", 0.5)

  i2_rr <- orchaRd::i2_ml(fit_rr$model)
  i2_vr <- orchaRd::i2_ml(fit_vr$model)

  dplyr::bind_rows(
    tibble::tibble(
      metric        = "lnRR",
      ri            = r,
      k             = fit_rr$model$k,
      estimate      = as.numeric(coef(fit_rr$model)),
      se            = fit_rr$model$se,
      ci.lb         = fit_rr$model$ci.lb,
      ci.ub         = fit_rr$model$ci.ub,
      tval          = fit_rr$model$zval,
      pval          = fit_rr$model$pval,
      sigma2_Study  = fit_rr$model$sigma2[1],
      sigma2_ES     = fit_rr$model$sigma2[2],
      sigma2_Strain = fit_rr$model$sigma2[3],
      I2_Total      = i2_rr[["I2_Total"]],
      I2_Study_ID   = i2_rr[["I2_Study_ID"]],
      I2_ES_ID      = i2_rr[["I2_ES_ID"]],
      I2_Strain     = i2_rr[["I2_Strain"]]
    ),
    tibble::tibble(
      metric        = "lnVR",
      ri            = r,
      k             = fit_vr$model$k,
      estimate      = as.numeric(coef(fit_vr$model)),
      se            = fit_vr$model$se,
      ci.lb         = fit_vr$model$ci.lb,
      ci.ub         = fit_vr$model$ci.ub,
      tval          = fit_vr$model$zval,
      pval          = fit_vr$model$pval,
      sigma2_Study  = fit_vr$model$sigma2[1],
      sigma2_ES     = fit_vr$model$sigma2[2],
      sigma2_Strain = fit_vr$model$sigma2[3],
      I2_Total      = i2_vr[["I2_Total"]],
      I2_Study_ID   = i2_vr[["I2_Study_ID"]],
      I2_ES_ID      = i2_vr[["I2_ES_ID"]],
      I2_Strain     = i2_vr[["I2_Strain"]]
    )
  )
})
```

```{r}
#| label: tbl_overall_ri
#| tbl-cap: "Overall lnRR and lnVR pooled estimates across the assumed within-subject correlation (ri), rho fixed at 0.5."

overall_sensitivity_ri %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed")) %>%
  scroll_box(width = "100%")
```

```{r}
#| label: save_overall_ri
readr::write_csv(overall_sensitivity_ri, here("..","data","sensitivity_ri_overall_results.csv"))
```

```{r}
#| label: contrast_ri
#| message: false
#| warning: false

contrast_sensitivity_ri <- purrr::map_dfr(ri_values, function(r) {

  dset  <- db_for_ri(r)
  db_vr <- dset %>% filter(!is.na(lnVR), !is.na(lnVR_var))

  k_counts    <- db_vr %>% count(Outcome_type)
  keep_levels <- k_counts %>% filter(n >= 5) %>% pull(Outcome_type)

  dat_f <- db_vr %>%
    filter(Outcome_type %in% keep_levels) %>%
    mutate(Outcome_type = droplevels(Outcome_type))

  VCV <- metafor::vcalc(
    vi = dat_f$lnVR_var, cluster = dat_f$Cohort_ID, obs = dat_f$ES_ID,
    rho = 0.5, data = dat_f
  )

  mod <- metafor::rma.mv(
    yi = dat_f$lnVR, V = VCV,
    mods = ~ Outcome_type,
    random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
    test = "t", method = "REML", sparse = TRUE, data = dat_f
  )

  cf <- coef(summary(mod))

  tibble::tibble(
    ri               = r,
    anxiety_est      = cf["intrcpt", "estimate"],
    anxiety_ci.lb    = cf["intrcpt", "ci.lb"],
    anxiety_ci.ub    = cf["intrcpt", "ci.ub"],
    anxiety_pval     = cf["intrcpt", "pval"],
    depression_est   = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"],
    contrast_est     = cf["Outcome_typeDepression", "estimate"],
    contrast_ci.lb   = cf["Outcome_typeDepression", "ci.lb"],
    contrast_ci.ub   = cf["Outcome_typeDepression", "ci.ub"],
    contrast_tval    = cf["Outcome_typeDepression", "tval"],
    contrast_pval    = cf["Outcome_typeDepression", "pval"]
  )
})
```

```{r}
#| label: tbl_contrast_ri
#| tbl-cap: "Depression-minus-anxiety lnVR contrast across the assumed within-subject correlation (ri), rho fixed at 0.5."

contrast_sensitivity_ri %>%
  mutate(across(where(is.numeric), ~round(., 4))) %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: save_contrast_ri
readr::write_csv(contrast_sensitivity_ri, here("..","data","sensitivity_ri_anxiety_depression_contrast.csv"))
```

```{r}
#| label: fig_overall_ri
#| fig-width: 9
#| fig-height: 5

p_sensitivity_ri <- overall_sensitivity_ri %>%
  mutate(metric = factor(metric, levels = c("lnRR", "lnVR"))) %>%
  ggplot(aes(x = factor(ri), y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = ci.lb, ymax = ci.ub), size = 0.6) +
  facet_wrap(~ metric, scales = "free_y") +
  labs(
    x = expression(paste("Assumed within-subject correlation, ", r[i])),
    y = "Pooled estimate (95% CI)",
    title = "Overall lnRR and lnVR estimates across assumed values of ri (rho = 0.5)"
  ) +
  theme_bw(base_size = 13)

p_sensitivity_ri
```

```{r}
#| label: save-figure_sensitivity_ri
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_ri_overall.pdf"),
       plot = p_sensitivity_ri,
       dpi = 300, device = cairo_pdf, width = 9, height = 5)

ggsave(filename = here("..","Plots", "sensitivity_ri_overall.jpg"),
       plot = p_sensitivity_ri,
       width = 9, height = 5, dpi = 300)
```

## Joint sensitivity to $r_i$ and $\rho$

`ri` and `rho` are separate assumptions -- one about the sampling variance of the 40 within-subject comparisons, the other about the covariance between different effect sizes sharing a cohort -- so we also cross them in a full 5 x 5 grid (`ri = 0.00, 0.25, 0.50, 0.75, 0.90` by `rho = 0.00, 0.25, 0.50, 0.75, 0.90`, 25 combinations in total) and refit the overall `lnRR`, overall `lnVR`, and the depression-minus-anxiety `lnVR` contrast at each combination, using the same model specification throughout.

```{r}
#| label: grid_ri_rho
#| message: false
#| warning: false

grid_ri_rho_results <- purrr::map_dfr(ri_values, function(r_i) {

  dset <- db_for_ri(r_i)

  purrr::map_dfr(rho_values, function(r_rho) {

    fit_rr <- fit_overall_mv(dset, "lnRR", "lnRR_var", r_rho)
    fit_vr <- fit_overall_mv(dset, "lnVR", "lnVR_var", r_rho)

    db_vr_g <- dset %>% filter(!is.na(lnVR), !is.na(lnVR_var))
    keep_levels_g <- db_vr_g %>% count(Outcome_type) %>% filter(n >= 5) %>% pull(Outcome_type)
    dat_fg <- db_vr_g %>% filter(Outcome_type %in% keep_levels_g) %>% mutate(Outcome_type = droplevels(Outcome_type))
    VCV_g <- metafor::vcalc(vi = dat_fg$lnVR_var, cluster = dat_fg$Cohort_ID, obs = dat_fg$ES_ID, rho = r_rho, data = dat_fg)
    mod_ct <- metafor::rma.mv(
      yi = dat_fg$lnVR, V = VCV_g, mods = ~ Outcome_type,
      random = list(~1 | Study_ID, ~1 | ES_ID, ~1 | Strain),
      test = "t", method = "REML", sparse = TRUE, data = dat_fg
    )
    cf <- coef(summary(mod_ct))

    dplyr::bind_rows(
      tibble::tibble(model = "lnRR_overall", ri = r_i, rho = r_rho, k = fit_rr$model$k,
                      estimate = as.numeric(coef(fit_rr$model)), se = fit_rr$model$se,
                      ci.lb = fit_rr$model$ci.lb, ci.ub = fit_rr$model$ci.ub,
                      tval = fit_rr$model$zval, pval = fit_rr$model$pval,
                      anxiety_est = NA_real_, depression_est = NA_real_),
      tibble::tibble(model = "lnVR_overall", ri = r_i, rho = r_rho, k = fit_vr$model$k,
                      estimate = as.numeric(coef(fit_vr$model)), se = fit_vr$model$se,
                      ci.lb = fit_vr$model$ci.lb, ci.ub = fit_vr$model$ci.ub,
                      tval = fit_vr$model$zval, pval = fit_vr$model$pval,
                      anxiety_est = NA_real_, depression_est = NA_real_),
      tibble::tibble(model = "lnVR_contrast_depression_minus_anxiety", ri = r_i, rho = r_rho, k = mod_ct$k,
                      estimate = cf["Outcome_typeDepression", "estimate"], se = cf["Outcome_typeDepression", "se"],
                      ci.lb = cf["Outcome_typeDepression", "ci.lb"], ci.ub = cf["Outcome_typeDepression", "ci.ub"],
                      tval = cf["Outcome_typeDepression", "tval"], pval = cf["Outcome_typeDepression", "pval"],
                      anxiety_est = cf["intrcpt", "estimate"],
                      depression_est = cf["intrcpt", "estimate"] + cf["Outcome_typeDepression", "estimate"])
    )
  })
})

readr::write_csv(grid_ri_rho_results, here("..","data","sensitivity_ri_rho_grid_results.csv"))
```

```{r}
#| label: tbl_grid_summary
#| tbl-cap: "Summary of the 25-combination ri x rho grid: range of pooled estimates and whether the 95% CI ever crosses zero (lnRR/contrast) or ever excludes zero (lnVR)."

grid_summary <- grid_ri_rho_results %>%
  group_by(model) %>%
  summarise(
    estimate_min   = round(min(estimate), 4),
    estimate_max   = round(max(estimate), 4),
    pval_min       = signif(min(pval), 3),
    pval_max       = signif(max(pval), 3),
    ci_crosses_zero    = any(ci.lb <= 0 & ci.ub >= 0),
    ci_excludes_zero_always = all(ci.lb > 0 | ci.ub < 0),
    sign_ever_flips = any(sign(estimate) != sign(estimate[1])),
    .groups = "drop"
  )

grid_summary %>%
  kable(caption = "Grid summary across all 25 ri x rho combinations.") %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))
```

```{r}
#| label: fig_grid_heatmap
#| fig-width: 10
#| fig-height: 4

model_labels <- c(
  lnRR_overall = "Overall lnRR",
  lnVR_overall = "Overall lnVR",
  lnVR_contrast_depression_minus_anxiety = "Depression - anxiety lnVR contrast"
)

p_grid <- grid_ri_rho_results %>%
  mutate(model_label = factor(model_labels[model], levels = model_labels)) %>%
  ggplot(aes(x = factor(ri), y = factor(rho), fill = estimate)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = round(estimate, 2)), size = 3) +
  facet_wrap(~ model_label, scales = "free") +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b", midpoint = 0) +
  labs(
    x = expression(paste("Assumed within-subject correlation, ", r[i])),
    y = expression(paste("Assumed within-cohort correlation, ", rho)),
    fill = "Estimate",
    title = "Pooled estimates across the joint ri x rho grid (25 combinations)"
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right")

p_grid
```

```{r}
#| label: save_fig_grid
#| eval: false

ggsave(filename = here("..","Plots", "sensitivity_ri_rho_grid.pdf"),
       plot = p_grid,
       dpi = 300, device = cairo_pdf, width = 10, height = 4)

ggsave(filename = here("..","Plots", "sensitivity_ri_rho_grid.jpg"),
       plot = p_grid,
       width = 10, height = 4, dpi = 300)
```

::: callout-note
```{r}
sessionInfo()
```
:::
