---
title: "CRIME-Q risk-of-bias figure"
---

```{r}
#| label: setup-crimeq
#| include: false
required_packages <- c(
  "dplyr",
  "ggplot2",
  "knitr",
  "readxl",
  "stringr",
  "tibble",
  "tidyr",
  "here"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]
if (length(missing_packages) > 0) {
  stop(
    "Missing required package(s): ",
    paste(missing_packages, collapse = ", ")
  )
}

invisible(lapply(required_packages, library, character.only = TRUE))
```

This chapter presents the finalized CRIME-Q music-exposure codebook and the study-level risk-of-bias rating figure. Both are read from the finalized Excel workbook so the displayed item definitions, item order, and plotted ratings stay aligned with the reviewed assessment.

```{r}
#| label: crimeq-load
#| include: false

input_file <- here("..", "data", "MUSIC-CRIME-Q_RoB_assessment.xlsx")
assessment_sheet <- "Sheet1"
codebook_sheet <- "CRIME-Q Codebook"

required_sheets <- c(assessment_sheet, codebook_sheet)
missing_sheets <- setdiff(required_sheets, excel_sheets(input_file))
if (length(missing_sheets) > 0) {
  stop(
    "The CRIME-Q workbook is missing required sheet(s): ",
    paste(missing_sheets, collapse = ", ")
  )
}

clean_text_cols <- function(.data) {
  .data |>
    rename_with(str_squish) |>
    mutate(across(everything(), ~ na_if(str_squish(.x), "")))
}

assessment_raw <- read_excel(
  path = input_file,
  sheet = assessment_sheet,
  col_types = "text",
  na = ""
) |>
  clean_text_cols()

codebook <- read_excel(
  path = input_file,
  sheet = codebook_sheet,
  col_types = "text",
  na = ""
) |>
  clean_text_cols() |>
  filter(!is.na(Item)) |>
  rename(
    Item_Name = `Item name`,
    Response_options = `Response options`,
    Decision_rule = `Decision rule`
  )

required_codebook_cols <- c(
  "Item", "Item_Name", "Question", "Response_options",
  "Yes", "Partly", "No", "Unclear", "NA", "Decision_rule"
)
if (!all(required_codebook_cols %in% names(codebook))) {
  stop("The CRIME-Q Codebook sheet does not contain the expected columns.")
}

item_cols <- codebook$Item

score_cols_in_workbook <- setdiff(names(assessment_raw), c("Study", "Study_Title"))
score_cols_in_workbook <- score_cols_in_workbook[
  !str_detect(score_cols_in_workbook, "_JUSTIFICATION$|_VERBATIM$")
]

if (!identical(score_cols_in_workbook, item_cols)) {
  stop("CRIME-Q score columns do not match the workbook codebook item order.")
}

assessment_matrix <- assessment_raw |>
  select(Study, Study_Title, all_of(item_cols))

if (any(is.na(assessment_matrix$Study))) {
  stop("Blank Study values detected in the CRIME-Q assessment sheet.")
}

if (anyDuplicated(assessment_matrix$Study) > 0) {
  stop("Duplicated Study values detected in the CRIME-Q assessment sheet.")
}

valid_ratings <- c("Yes", "Partly", "No", "Unclear", "NA")
observed_ratings <- assessment_matrix |>
  select(all_of(item_cols)) |>
  unlist(use.names = FALSE) |>
  unique() |>
  na.omit()

unexpected_ratings <- setdiff(observed_ratings, valid_ratings)
if (length(unexpected_ratings) > 0) {
  stop(
    "Unexpected CRIME-Q rating values found: ",
    paste(unexpected_ratings, collapse = ", ")
  )
}

study_label <- function(study_id) {
  str_replace(study_id, "^([[:alpha:]]+)([0-9]{4})$", "\\1 et al. \\2")
}

assessment_matrix <- assessment_matrix |>
  mutate(
    Study_Label = study_label(Study),
    .after = Study
  )
```

## CRIME-Q item definitions

The table below is read from the `CRIME-Q Codebook` sheet in the assessment workbook. It defines each item, the allowed response options, and the item-specific scoring rules used for the figure.

```{r}
#| label: crimeq-definitions
#| echo: false

codebook |>
  transmute(
    Item,
    `Item name` = Item_Name,
    Question,
    `Response options` = Response_options,
    Yes,
    Partly,
    No,
    Unclear,
    `NA`,
    `Decision rule` = Decision_rule
  ) |>
  kable(align = c("l", "l", "l", "l", "l", "l", "l", "l", "l", "l"))
```

## CRIME-Q study-level rating figure

The source assessment contains `r nrow(assessment_matrix)` study rows and `r length(item_cols)` CRIME-Q items. Ratings are plotted directly from the finalized study-level score cells in `Sheet1`; justification and verbatim columns are retained in the workbook but are not plotted.

```{r}
#| label: crimeq-heatmap
#| fig-width: 13
#| fig-height: 8
#| out-width: "100%"
#| dev.args: !expr list(bg = "transparent")
#| fig-cap: "CRIME-Q item ratings from the finalized study-level assessment. Green = Yes, yellow = Partly, red = No, orange = Unclear, and blue = NA."

plot_dat <- assessment_matrix |>
  pivot_longer(
    cols = all_of(item_cols),
    names_to = "Item",
    values_to = "Rating"
  ) |>
  mutate(
    Item = factor(Item, levels = item_cols),
    Study_Label = factor(Study_Label, levels = rev(assessment_matrix$Study_Label)),
    Rating = factor(Rating, levels = valid_ratings)
  )

rating_colors <- c(
  "Yes" = "#009E60",
  "Partly" = "#EFBF04",
  "No" = "#DC143C",
  "Unclear" = "#F4B183",
  "NA" = "#D9EAF7"
)
gallery_text <- "#242321"

CRIME <- ggplot(plot_dat, aes(x = Item, y = Study_Label, fill = Rating)) +
  geom_tile(color = "#FFFDF8", linewidth = 0.6, width = 0.96, height = 0.96) +
  scale_fill_manual(
    values = rating_colors,
    breaks = valid_ratings,
    drop = FALSE,
    name = NULL
  ) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 22)) +
  coord_equal() +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    legend.key = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    axis.text.x = element_text(
      hjust = 0.5,
      vjust = 0,
      face = "bold",
      color = gallery_text
    ),
    axis.text.y = element_text(face = "bold", color = gallery_text),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 11, color = gallery_text),
    plot.margin = margin(10, 10, 10, 10)
  )

CRIME
```

```{r}
#| label: save-figure_crimeq
#| eval: false

ggsave(
  filename = here("..", "Plots", "CRIMEQ.pdf"),
  plot = CRIME,
  dpi = 300,
  device = cairo_pdf,
  width = 12,
  height = 8
)

ggsave(
  filename = here("..", "Plots", "CRIMEQ.jpg"),
  plot = CRIME,
  width = 12,
  height = 8,
  dpi = 300
)
```
