---
title: "CRIME-Q risk-of-bias figure"
---

```{r}
#| label: setup-crimeq
#| include: false
pacman::p_load(
  dplyr,
  ggplot2,
  knitr,
  readr,
  readxl,
  stringr,
  tibble,
  tidyr,
  here
)
```

This chapter reproduces the CRIME-Q study-level rating figure from the cohort-level risk-of-bias assessment table and provides a compact item definition section for the CRIME-Q music-exposure codebook.

```{r}
#| label: crimeq-load
#| include: false

input_file <- here("..", "data", "MUSIC-CRIME-Q_RoB_assessment.xlsx")
input_sheet <- "Table 2"
codebook_file <- here("..", "CRIME-Q_Music_RoB_Codebook_v1.8.csv")

item_cols <- c(
  "1X", "2X", "2Y", "3X", "3Y", "3Z", "4Y",
  "5X", "5Y", "5Z(1)", "5Z(2)", "5Z(3)",
  "6X", "7X", "7Z(1)", "7Z(2)", "7Z(3)",
  "8X", "8Z(1)", "8Z(2)", "9X", "10X", "10Z"
)

expected_cols <- c("Study_ID", "Cohort_ID", item_cols)

cohort_matrix <- read_excel(
  path = input_file,
  sheet = input_sheet,
  col_types = "text",
  na = ""
) |>
  rename_with(str_trim) |>
  mutate(across(everything(), str_trim)) |>
  select(all_of(expected_cols))

if (!identical(names(cohort_matrix), expected_cols)) {
  stop("Sheet 'Table 2' does not match the expected column structure.")
}

if (anyDuplicated(cohort_matrix$Cohort_ID) > 0) {
  stop("Duplicated Cohort_ID values detected.")
}

collapse_rating <- function(x) {
  vals <- unique(x)
  if (length(vals) == 1L) vals else "Partly"
}

study_matrix <- cohort_matrix |>
  group_by(Study_ID) |>
  summarise(
    across(all_of(item_cols), collapse_rating),
    .groups = "drop"
  ) |>
  arrange(Study_ID) |>
  mutate(ID = row_number()) |>
  relocate(ID, .before = Study_ID)

codebook <- read_csv(
  codebook_file,
  show_col_types = FALSE,
  trim_ws = TRUE
) |>
  arrange(Order)

if (!identical(codebook$Item, item_cols)) {
  stop("Codebook item order does not match the CRIME-Q matrix columns.")
}
```

## CRIME-Q study-level rating matrix

The source assessment contains `r nrow(cohort_matrix)` cohort rows from `r n_distinct(cohort_matrix$Study_ID)` studies. When a study has multiple cohorts, item ratings are collapsed to one display row: if all cohorts have the same rating, that rating is retained; if cohort ratings differ, the study is displayed as `Partly`.

```{r}
#| label: crimeq-heatmap
#| fig-width: 14
#| fig-height: 7
#| out-width: "100%"
#| fig-cap: "CRIME-Q item ratings collapsed to one row per study. Green = Yes, yellow = Partly or Unclear, red = No, and blue = NA."

plot_dat <- study_matrix |>
  pivot_longer(
    cols = all_of(item_cols),
    names_to = "Item",
    values_to = "Rating"
  ) |>
  mutate(
    Item = factor(Item, levels = item_cols),
    ID = factor(ID, levels = rev(unique(ID))),
    Fill_group = case_when(
      Rating == "Yes" ~ "Yes",
      Rating %in% c("Partly", "Unclear") ~ "Partly/Unclear",
      Rating == "No" ~ "No",
      Rating == "NA" ~ "NA",
      TRUE ~ "Other"
    )
  )

if (any(plot_dat$Fill_group == "Other")) {
  warning("Some unexpected CRIME-Q rating values were found.")
}

ggplot(plot_dat, aes(x = Item, y = ID, fill = Fill_group)) +
  geom_tile(color = "black", linewidth = 0.35) +
  scale_fill_manual(
    values = c(
      "Yes" = "#33FF33",
      "Partly/Unclear" = "#FFFF4D",
      "No" = "#FF3333",
      "NA" = "#1E3DFF"
    ),
    breaks = c("Yes", "Partly/Unclear", "No", "NA"),
    name = NULL
  ) +
  labs(
    x = NULL,
    y = "ID"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(
      angle = 0,
      hjust = 0.5,
      vjust = 0.5,
      face = "bold"
    ),
    axis.text.y = element_text(face = "bold"),
    axis.ticks = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 11),
    plot.margin = margin(10, 10, 10, 10)
  )
```

## Study ID lookup

```{r}
#| label: crimeq-id-lookup
#| echo: false

study_matrix |>
  select(ID, Study_ID) |>
  kable(
    col.names = c("Figure ID", "Study ID"),
    align = c("r", "l")
  )
```

## CRIME-Q item definitions

The compact guide below keeps the definition fields needed to interpret the figure: item code, construct, assessment level, allowed ratings, and the simplified coding rule.

```{r}
#| label: crimeq-codebook-summary
#| echo: false

construct_labels <- c(
  QoR = "Quality of reporting",
  MQ = "Methodological quality",
  RoB = "Risk of bias"
)

codebook |>
  transmute(
    Item,
    Construct = recode(Construct, !!!construct_labels),
    Domain,
    `Assessment level` = Assessment_level,
    `Ratings` = Response_options
  ) |>
  kable(align = c("l", "l", "l", "l", "l"))
```

```{r}
#| label: crimeq-item-definitions
#| echo: false
#| results: asis

clean_field <- function(x) {
  x <- str_squish(as.character(x))
  ifelse(is.na(x) | x == "" | x == "-", NA_character_, x)
}

for (i in seq_len(nrow(codebook))) {
  row <- codebook[i, ]
  construct <- recode(row$Construct, !!!construct_labels)
  syrcle <- clean_field(row$SYRCLE_item)
  na_rule <- clean_field(row$NA_rule)
  notes <- clean_field(row$Notes)

  cat("### ", row$Item, ": ", row$Domain, "\n\n", sep = "")
  cat("**Construct:** ", construct, "  \n", sep = "")
  if (!is.na(syrcle)) {
    cat("**SYRCLE mapping:** ", syrcle, "  \n", sep = "")
  }
  cat("**Assessment level:** ", row$Assessment_level, "  \n", sep = "")
  cat("**Rating options:** ", row$Response_options, "\n\n", sep = "")
  cat("**What it checks:** ", row$Question, "\n\n", sep = "")
  cat("**Simplified coding rule:** ", row$Operational_guidance, "\n\n", sep = "")
  if (!is.na(na_rule) && na_rule != "Not used") {
    cat("**NA rule:** ", na_rule, "\n\n", sep = "")
  }
  if (!is.na(notes)) {
    cat("**Note:** ", notes, "\n\n", sep = "")
  }
}
```
